Neural and Symbolic Spatiotemporal Optimization of Cosmological Particle-Mesh Simulation

### 2023-05-17 docs/papers/sto/sto.tex
FIXUP draft: correction and comments

A mean squared error (MSE) loss on the fields is dominated by the small
scale modes.
In this work we choose uniform weights in the scale space on the
relative MSE.
We bin the modes logarithmically by \nicefrac13 octave in scale, e.g.,
in the $i$-th $k$-bin, $K_i \triangleq [k_i, k_{i+1})$, with $k_i =
2^{1+i/3} \pi / L$ and $i = 0, 1, \cdots$.
Our loss on $f$ sums up the MSE relative to the target field power in
logarithmic bins:
%
\begin{equation}
\cL_f = \sum_i
\frac{\sum_{k \in K_i} \bigl| \hat f(\vk) - f(\vk) \bigr|^2}
     {\sum_{k' \in K_i} |f(\vk')|^2}.
\end{equation}
%
To understand the form of the loss function, let's introduce for each
mode the square root ratio of power
%
\begin{equation}
T(\vk) \triangleq
\sqrt{\frac{|\hat f(\vk)|^2}{|f(\vk')|^2}},
\end{equation}
%
and the cross correlation coefficient
%
\begin{equation}
r(\vk) \triangleq
\frac{\Re \bigl[ \hat f(\vk) f^*(\vk) \bigr]}
     {\sqrt{|\hat f(\vk)|^2 |f(\vk')|^2}}.
\end{equation}
%
Now, under the mild assumptions that $T$ and $r$ are smooth and
isotropic, we can show
%
\begin{equation}
\cL_f \simeq \sum_\vk \Bigl\{ \bigl[ 1 - T(\vk) \bigr]^2
  + 2 T(\vk) \bigl[ 1 - r(\vk) \bigr] \Bigr\} \geq 0,
\end{equation}
%
that only vanishes when both $T$ and $r$ are unity for all modes, i.e.,
when $\hat f$ is identical to $f$.
Above we choose the binned version instead for the robustness against
the noisy modewise $|f(\vk)|^2$ in the denominator.

@@

The curvature $\Omega_k$ is related to the separate universe simulation
\citep{LiEtAl2014a, WagnerEtAl2015}.

>>>>

A mean squared error (MSE) loss on the fields is dominated by the small
scale modes \citep{HeEtAl2019}.
In this work we choose uniform weights in the scale space on the
relative MSE.
We bin the modes logarithmically by \nicefrac13 octave in scale, with
the $i$-th $k$-bin $K_i \triangleq [k_i, k_{i+1})$, where $k_i =
2^{1+i/3} \pi / L$.
Our loss on $f$ sums up the MSE relative to the target field power in
logarithmic bins:
%
\begin{equation}
\cL_f = \sum_i
\frac{\sum_{k \in K_i} \bigl| \hat f(\vk) - f(\vk) \bigr|^2}
     {\sum_{k' \in K_i} |f(\vk')|^2}.
\end{equation}
%
To understand the form of the loss function, let's introduce for each
mode the square root ratio of power
%
\begin{equation}
T(\vk) \triangleq
\sqrt{\frac{|\hat f(\vk)|^2}{|f(\vk')|^2}},
\end{equation}
%
and the cross correlation coefficient
%
\begin{equation}
r(\vk) \triangleq
\frac{\Re \bigl[ \hat f(\vk) f^*(\vk) \bigr]}
     {\sqrt{|\hat f(\vk)|^2 |f(\vk')|^2}}.
\end{equation}
%
Now, under the mild assumptions that $T$ and $r$ are smooth and
isotropic, we can show
%
\begin{equation}
\cL_f \propto \int \!\d\ln\!k\, \Bigl\{ \bigl[ 1 - T(\vk) \bigr]^2
  + 2 T(\vk) \bigl[ 1 - r(\vk) \bigr] \Bigr\} \geq 0,
\end{equation}
%
that only vanishes when both $T$ and $r$ are unity for all modes, i.e.,
when $\hat f$ is identical to $f$.
Above we choose the binned version instead for the robustness against
the noisy modewise $|f(\vk)|^2$ in the denominator.
The derivation here is similar to that in \citet{HeEtAl2019} on the MSE
loss, while our modification is important to balance among all scales so
that the neural network does not focus too much on either high or low
$k$'s.

@@

The curvature $\Omega_k$ is related to the separate universe simulation
\citep{LiEtAl2014}.
